---
title: "Simulating Probabilities & Datasets"
format: 
  revealjs:
      theme: [simple, ../style.scss]
      auto-stretch: false
editor: source
execute:
  echo: true
---

```{r setup}
#| include: false
#| message: false
library(tidyverse)

set.seed(435)

fake_data <- tibble(names   = charlatan::ch_name(1000),
        height  = rnorm(1000, mean = 67, sd = 3),
        age     = runif(1000, min = 15, max = 75),
        measure = rbinom(1000, size = 1, prob = 0.6)) |> 
  mutate(supports_measure_A = ifelse(measure == 1, "yes", "no"))

```

Today we will...

-   New Material
    -   Thinking through a probability simulation task
    -   Four Steps to Simulate Datasets
    -   In-line R Code
    -   Survey on Experiences in STAT 331 / 531
-   Work Time
    -   Lab 9
    -   Challenge 9


# The Birthday Problem

> What the probability that at least two people in a group of randomly selected
people share a birthday?

## Our Task

Simulate the approximate probability that at least two people have the same
birthday (same day of the year, not considering year of birth or leap years).


## Writing a function to...

+ simulate the birthdays of 50 people.
+ count how many birthdays are shared.
+ return whether or not a shared birthday exists.


## Step 1 -- simulate the birthdays of 50 people

```{r}
#| code-line-numbers: false

get_shared_birthdays <- function(n = 50){
  
  bday_data <- tibble(person = 1:n,
                      bday   = sample(1:365, 
                                      size = n, 
                                      replace = TRUE)
                      )
}
```

. . .

::: {.callout-tip}
# Why did I set `replace = TRUE` for the `sample()` function?
:::

## Step 2 -- count how many birthdays are shared

```{r}
#| code-line-numbers: "9-12"

get_shared_birthdays <- function(n = 50){
  
  bday_data <- tibble(person = 1:n,
                      bday   = sample(1:365, 
                                      size = n, 
                                      replace = TRUE)
                      )
  
  double_bdays <- bday_data |> 
    count(bday) |> 
    filter(n >= 2) |> 
    nrow()
  
}
```

## Step 3 -- return whether or not a shared birthday exists

::: {.small}
```{r}
#| code-line-numbers: "13"

get_shared_birthdays <- function(n = 50){
  bday_data <- tibble(person = 1:n,
                      bday   = sample(1:365, 
                                      size = n, 
                                      replace = T)
                      )
  
  double_bdays <- bday_data |> 
    count(bday) |> 
    filter(n >= 2) |> 
    nrow()
  
  return(double_bdays > 0)
}
```
:::

. . .

::: {.callout-tip}
# What type of output is this function returning? 
:::

## Using Function to Simulate Many Probabilities

Use a `map()` function to simulate 1000 datasets.

::: {.midi}
```{r}
#| code-line-numbers: false
sim_results <- map_lgl(.x = 1:1000,
                       .f = ~ get_shared_birthdays(n = 50))

```
:::

. . .

</br>

+ What proportion of these datasets contain at least two people with the same
birthday?

::: {.midi}
```{r}
#| code-line-numbers: false

sum(sim_results) / 1000
```
:::



# Four Steps to Simualate Datasets

## Simulate Multiple Datasets - Step 1

::: {.small}
Write a function to simulate height data from a population with some mean and SD
height. 

The user should be able to input:

- how many observations to simulate
- the mean and standard deviation of the Normal distribution to use when 
simulating

::: {.callout-tip}
## Pencil out your function! 
:::
:::



::: columns
::: {.column width="60%"}
::: {.fragment}
::: {.small}
```{r}
#| code-line-numbers: false

sim_ht <- function(n = 200, avg, std){
  tibble(person = 1:n,
         ht = rnorm(n = n, mean = avg, sd = std))
}
```
:::
:::
:::

::: {.column width="5%"}
:::

::: {.column width="35%"}
::: {.fragment}
::: {.small}
```{r}
#| code-line-numbers: false

sim_ht(n = 5, 
        avg = 66, 
        std = 3)
```
:::
:::
:::
:::



## Simulate Multiple Datasets - Step 2

Create a set of parameters (mean and SD) for each population.

```{r}
#| code-line-numbers: false

crossing(mean_ht = seq(from = 60, to = 78, by = 6),
            std_ht  = c(3, 6))
```

## Simulate Multiple Datasets - Step 3

Simulate datasets with different mean and SD heights.

::: {.small}
```{r}
#| code-line-numbers: "4-7"

crossing(mean_ht = seq(from = 60, to = 78, by = 6),
         std_ht  = c(3, 6)
         ) |> 
 mutate(ht_data = pmap(.l = list(avg = mean_ht, std = std_ht), 
                       .f = sim_ht
                       )
        )
```
:::

. . .

::: {.small}
::: {.callout-tip}
# Why am I getting a tibble in the `ht_data` column?
:::
:::

## Simulate Multiple Datasets - Step 4

Extract the contents of each list!

::: {.small}
```{r}
#| code-line-numbers: "8"
#| eval: false

crossing(mean_ht = seq(from = 60, to = 78, by = 6),
         std_ht  = c(3, 6)
         ) |> 
 mutate(ht_data = pmap(.l = list(avg = mean_ht, std = std_ht), 
                       .f = sim_ht
                       )
        ) |> 
  unnest(cols = ht_data) |> 
  slice_sample(n = 10)
```
:::

::: columns
::: {.column width="50%"}
::: {.small}
```{r}
#| echo: false
#| eval: true

crossing(mean_ht = seq(from = 60, to = 78, by = 6),
         std_ht  = c(3, 6)
         ) |> 
 mutate(ht_data = pmap(.l = list(avg = mean_ht, std = std_ht), 
                       .f = sim_ht
                       )
        ) |> 
  unnest(cols = ht_data) |> 
  slice_head(n = 10)
```
:::
:::

::: {.column width="5%"}
:::

::: {.column width="45%"}
::: {.midi}
::: {.fragment}
Why do I now have `person` and `ht` columns?
:::

::: {.fragment}
How many rows should I have for each `mean_ht`, `std_ht` combo?
:::
:::
:::
:::

## `nest()` and `unnest()`

-   We can pair functions from the `map()` family very nicely with two `tidyr`
functions: `nest()` and `unnest()`.
-   These allow us to easily map functions onto subsets of the data.

## `nest()`

**Nest** subsets of the data (as tibbles) inside a tibble.

-   Specify the column(s) to create subsets on.

::: columns
::: {.column width="50%"}
```{r}
#| echo: true
#| code-line-numbers: false
#| eval: false

mtcars |> 
  nest(.by = cyl)
```
:::

::: {.column width="50%"}
```{r}
#| code-line-numbers: false
#| echo: false

mtcars |> 
  nest(.by = cyl)
```
:::
:::

## `unnest()`

**Un-nest** the data by row binding the subsets back together.

-   Specify the column(s) that contains the subsets.

::: columns
::: {.column width="40%"}
::: {.small}
```{r}
#| echo: true
#| eval: false
#| code-line-numbers: false

mtcars |> 
  nest(.by = cyl) |> 
  unnest(data) |> 
  slice_head(n = 10)
```
:::
:::

::: {.column width="5%"}
:::

::: {.column width="55%"}
::: {.small}
```{r}
#| echo: false
#| code-line-numbers: false

mtcars |> 
  nest(.by = cyl) |> 
  unnest(data) |> 
  slice_head(n = 10)
```
:::
:::
:::

## Simulate Multiple Datasets - Step 5

Plot the samples simulated from each population.

::: {.small}
```{r}
#| echo: false
#| code-line-numbers: false

fake_ht_data <- crossing(mean_ht = seq(from = 60, to = 78, by = 6),
         std_ht  = c(3, 6)
         ) |> 
 mutate(ht_data = pmap(.l = list(avg = mean_ht, std = std_ht), 
                       .f = sim_ht
                       )
        ) |> 
  unnest(cols = ht_data)
```

```{r}
#| code-fold: true
#| fig-width: 8
#| fig-height: 4
#| fig-align: center

fake_ht_data |> 
  mutate(across(.cols = mean_ht:std_ht, 
                .fns = ~as.character(.x)), 
         mean_ht = fct_recode(mean_ht, 
                              `Mean = 60` = "60", 
                              `Mean = 66` = "66", 
                              `Mean = 72` = "72", 
                              `Mean = 78` = "78"), 
         std_ht = fct_recode(std_ht, 
                             `Std = 3` = "3", 
                             `Std = 6` = "6")
         ) |> 
  ggplot(mapping = aes(x = ht)) +
  geom_histogram(color = "white") +
  facet_grid(std_ht ~ mean_ht) +
  labs(x = "Height (in)",
       y = "",
       subtitle = "Frequency of Observations",
       title = "Simulated Heights from Eight Different Populations")
```
:::

## In-line Code

We can automatically include code output in the written portion of a Quarto
document using `` `r` ``.

. . .

</br> 

```{r}
#| code-line-numbers: false

my_rand <- rnorm(1, mean = 0, sd = 1)
my_rand
```

</br> 

Type this: My random number is `` `r knitr::inline_expr("my_rand")` ``.

To get this: My random number is `r my_rand`.

# [Lab 9: Data Simulation Exploration](../../labs/instructions/lab-9-instructions.qmd)

- One **required** probability simulation
- One *optional* dataset simulation

# [Challenge 9: Formatting Nice Tables](../../labs/instructions/challenge-9-instructions.qmd)

- Change tables from Lab 8 using `kable()`
- Change table(s) from Lab 9 using `gt()`

# [Survey on Experiences in STAT 331 / 531](https://forms.gle/wtXGQTFq5yzrX32W6)

::: {.midi}
Anonymous Google Form: <https://forms.gle/wtXGQTFq5yzrX32W6>
:::

. . .

::: {.midi}
> If we get an 85% completion rate I will bring pizza to class next Thursday. 
:::