---
title: "Debugging Errors!"
format: revealjs
---

```{r}
#| label: setup
#| echo: false
#| message: false
#| warning: false

library(tidyverse)
library(flair)
```

# Syntax Errors

## Did you leave off a parenthesis?

</br>

seq[(]{style="background-color: #ffff7f"}from = 1, to = 10, by = 1

```{r}
#| echo: true
#| error: true
#| code-line-numbers: false
seq(from = 1, to = 10, by = 1
```

. . .

</br>

```{r}
#| echo: true
#| code-line-numbers: false
seq(from = 1, to = 10, by = 1)
```

## Did you leave off a comma?

</br>

seq(from = 1, to = 10 [by]{style="background-color:#ffff7f"} = 1)

```{r}
#| echo: true
#| error: true
#| code-line-numbers: false
seq(from = 1, to = 10 by = 1)
```

. . .

</br>

```{r}
#| echo: true
#| code-line-numbers: false
seq(from = 1, to = 10, by = 1)
```

## Are you using the right function name?

</br>

[sequence]{style="background-color:#ffff7f"}(from = 1, to = 10, by = 1)

```{r}
#| echo: true
#| error: true
#| code-line-numbers: false
sequence(from = 1, to = 10, by = 1)
```

. . .

</br>

```{r}
#| echo: true
#| code-line-numbers: false

seq(from = 1, to = 10, by = 1)
```

# Object Type Errors

## Are you using the right *input* that the function expects?

</br>

sqrt(["1"]{style="background-color:#ffff7f"})

```{r}
#| echo: true
#| error: true
#| code-line-numbers: false

sqrt("1")
```

. . .

</br>

```{r}
#| echo: true
#| code-line-numbers: false

sqrt(1)
```

## Are you expecting the right *output* of the function?

</br>

my_obj[(5)]{style="background-color:#ffff7f"}

```{r}
#| echo: true
#| error: true
#| code-line-numbers: false

my_obj <- seq(from = 1, to = 10, by = 1)

my_obj(5)
```

. . .

</br>

```{r}
#| echo: true
#| code-line-numbers: false

my_obj[5]
```

# Errors + Warnings + Messages

## Messages

Just because you see scary red text, this does **not** mean something went wrong! This is just R communicating with you.

. . .

</br>

For example, you will often see:

```{r}
#| echo: true
#| message: true
#| warning: true
#| code-line-numbers: false

library(lme4)
```

## Warnings

Often, R will give you a **warning**.

-   This means that your code *did* run...

-   ...but you probably want to make sure it succeeded.

. . .

</br>

::: columns
::: {.column width="50%"}
```{r}
#| echo: true
#| warning: true
#| code-line-numbers: false

my_vec <- c("a", "b", "c")

my_new_vec <- as.integer(my_vec)
```
:::

::: {.column width="5%"}
:::

::: {.column width="45%"}
::: fragment
**Does this look right?**

```{r}
#| echo: true
#| code-line-numbers: false

my_new_vec
```
:::
:::
:::

## Errors

If the word **Error** appears in your message from R, then you have a problem.

This means your code **could not run**!

. . .

</br>

```{r}
#| echo: true
#| error: true
#| code-line-numbers: false
my_vec <- c("a", "b", "c")

my_new_vec <- my_vec + 1
```

# Parlez-vous ERROR?

## **R** says...

> Error: Object `some_obj` not found.

. . .

It *probably* means...

> You haven't run the code to create `some_obj` OR you have a typo in the name!

. . .

</br>

```{r}
#| echo: true
#| error: true
#| code-line-numbers: false
some_ojb <- 1:10

mean(some_obj)
```

## **R** says...

> Error: Object of type 'closure' is not subsettable.

. . .

It *probably* means...

> Oops, you tried to use square brackets on a function.

. . .

</br>

```{r}
#| echo: true
#| error: true
#| code-line-numbers: false

mean[1, 2]
```

## **R** says...

> Error: Non-numeric argument to binary operator.

. . .

It *probably* means...

> You tried to do math on data that isn't numeric.

. . .

</br>

```{r}
#| echo: true
#| error: true
#| code-line-numbers: false
"a" + 2
```

## What if none of these solved my error?

1.  Look at the help file for the function! (e.g., `?group_by`)

2.  When all else fails, Google your error message or ask ChatGPT!

-   Leave out the specifics.

-   Include the name(s) of the function(s) you are using.

## Try it...

**What's wrong here?**

```{r}
#| echo: true
#| error: true
#| code-line-numbers: false

matrix(
  c("a", "b", "c", "d"), 
  num_row = 2
  )
```

. . .

</br>

The documentation says...

```         
matrix(data = NA, nrow = 1, ncol = 1, byrow = FALSE,
       dimnames = NULL)
```
