---
title: "An Introduction to the tidyverse & dplyr"
format: revealjs
---

# R History lesson
# What is the `tidyverse`?

## The tidyverse

The `tidyverse` is a collection of packages that follow a (somewhat) consistent philosophy and structure.

The term `base R` is sometimes used as an alternative, to refer to only R's built-in packages.

## A brief history

. . .

**2008:** `plyr` and `ggplot` released as part of **Hadley Wickham**'s doctoral dissertation.

. . .

**2012:** The pipe (`%>%`) is created.  

*(While the pipe itself is not technically tidyverse, the tidyverse is built around its use.)*

. . .

**2014:** `dplyr` released; Wickham hired as Chief Scientist at **RStudio**

. . .

<!-- Add when native pipe was added -->

**Today:** A paid team of 20-ish developers maintains a huge number of packages.  (https://www.tidyverse.org/)


## The Philosophy of the tidyverse

- Keep functions as simple as possible 
  * Each function should do one thing well 

. . .

- Function names should be verbs
  * When many functions use the same verb, focus on noun (ggplot2)
  * Favor explicit, lengthy names, over short, implicit, names

. . .

- Build **pipe-friendly** functions

##

> Programs must be written for people to read, and only incidentally for 
> machines to execute.
>
> — Hal Abelson

</br>

```{r}
#| eval: false

penguins %>%
  filter(species == "Adelie")
```

```{r}
#| eval: false

penguins[penguins$species == "Adelie",]
```

## The tidyverse: Pros 

* Fairly consistent behavior of functions

    + Predictable **input** and **output** types
    + *"opinionated"*
    + More on this later
    
. . .

* Semi-guaranteed maintenance of packages (because RStudio pays developers!)

. . .

* More available and polished documentation and tutorials

. . .

* Welcoming community


## The tidyverse: Cons

::: {.incremental}
* Sometimes less flexible for complex tasks

* Generally slower, especially on big data

* Updates frequently -> sometimes this makes your old code unrunnable.

* Some people object to **RStudio**'s strong influence over an open-source community.
:::


## Should you use the tidyverse???

. . .

Even the most die-hard tidyverse fanatics use base R and other packages quite frequently.

. . .

Above all, the best code approach is the one you enjoy the most and that gets the job done!

. . .


In this class, you must demonstrate an understanding of the packages we have 
learned. 

. . .

However, after this class you are free to use whatever package(s) you want!
