---
title: "PA 3: Identify the Mystery College 🏫"
subtitle: "Data Wrangling with `dplyr`"
author: "Names of Your Group!"
format: html
embed-resources: true
code-tools: true
toc: true
editor: source
execute: 
  error: true
  echo: true
  eval: false
  message: false
  warning: false
---

Today you will use the `dplyr` package to clean some data. We will then use that
cleaned data to figure out what college Margaret has been accepted to.

***This task is complex. It requires many different types of abilities. Everyone will be good at some of these abilities but nobody will be good at all of them. In order to produce the best product possible, you will need to use the skills of each member of your group.***

## Groupwork Protocols

During the Practice Activity, you and your partner will alternate between two
roles---Developer and Coder. 

When you are the **Developer**, you will type into the Quarto document in
RStudio. However, you **do not** type your own ideas. Instead, you type what 
the Coder tells you to type. You are permitted to ask the Coder clarifying
questions, and, if both of you have a question, you are permitted to ask the
professor. You are expected to run the code provided by the Coder and, if 
necessary, to work with the Coder to debug the code. Once the code runs, you
are expected to collaborate with the Coder to write code comments that describe
the actions taken by your code.

When you are the **Coder**, you are responsible for reading the 
instructions / prompts and directing the Developer what to type in the Quarto
document. You are responsible for managing the resources your group has
available to you (e.g., cheatsheet, textbook). If necessary, you should work
with the Developer to debug the code you specified. Once the code runs, you
are expected to collaborate with the Developer to write code comments that
describe the actions taken by your code.

### Group Norms

Remember, your group is expected to adhere to the following norms:

1.  Think and work together. Do not divide the work.
2.  You are smarter together.
3.  No cross-talk with other groups.

# Part One: Data Import and Cleaning

This section will clean today's data so that you can use it more easily in Part
Two.

The partner whose birthday comes last starts as the Developer (typing and
listening to instructions from the Coder)!

## Data Download & Package Loading

First, we declare our package dependencies and load the data.

*The data loading function `read_csv()` will give you an outpouring of helpful information about the dataset. If you do not see the word "error", there is nothing to be concerned about.*

```{r}
#| warning: false
#| message: false
#| label: load-data-packages

library(tidyverse)

colleges <- read_csv("https://www.dropbox.com/s/bt5hvctdevhbq6j/colleges.csv?dl=1")
```

Take a look at the variables in your downloaded data by running the following
code. This tells reports the data type for each column in the dataset.

```{r}
glimpse(colleges)
```

## Data Cleaning

Now we will clean the data. Alas, each of the R chunks in this section will
cause an error and / or do the desired task incorrectly. Even the chunks that
run without error **are not** correct! You will need to find the mistake and
correct it to complete the intended action.

**Step 1:** There are too many variables in this data set. We don't need all of
them. Narrow your data set down to only:

-   `INSTNM` name of the institution
-   `CITY` city, `STABBR` state, and `ZIP` ZIP code of the institution
-   `ADM_RATE` admissions rate
-   `SAT_AVG` average SAT score
-   `UGDS` number of undergraduate students
-    `TUITIONFEE_IN` in- and `TUITIONFEE_OUT` out-of-state tuition
-   `CONTROL` Whether the school is public or private
-   `REGION` region of the school.

```{r}
#| label:  select-vars
colleges_clean1 <- colleges | >  
  select(INSTNM, CITY, STABBR, ZIP,
         ADM_RATE, SAT_AVG, UGDS,
         TUITIONFEE_IN, TUITIONFEE_OUT
         CONTROL, REGION) 
```

**Step 2:** Remove the schools that are for-profit (category 3), keeping public
(category 1) and private schools (category 2).

```{r}
#| label: filter-private-profit

colleges_clean2 <- colleges_clean |> 
  filter(CONTROL == 1, CONTROL == 2)
```

**Step 3:** Adjust the appropriate variables to be numeric, using `as.numeric()`.

```{r}
#| label: mutate-numeric

colleges_clean3 <- colleges_clean |> 
  mutate(TUITIONFEE_IN  = numeric(TUITIONFEE_IN),
         TUITIONFEE_OUT = numeric(TUITIONFEE_OUT),
         SAT_AVG        = numeric(SAT_AVG),
         UGDS           = numeric(UGDS),
         ADM_RATE       = numeric(ADM_RATE)) 
```

<!-- Swap roles -- Developer becomes Coder, Coder becomes Developer! -->

**Step 4:** Adjust the appropriate variables to be factors, using `as.factor()`.

```{r}
#| label: mutate-factor

colleges_clean4 <- colleges_clean |>
  mutate(CONTROL = as.character(CONTROL),
         REGION  = as.character(REGION))
```

**Step 5:** Create a new variable called `TUITION_DIFF` which contains the
difference between out-of-state and in-state costs.

```{r}
#| label: add-differences-between-in-state-out-of-state

colleges_clean5 |> 
    TUITION_DIFF = TUITIONFEE_OUT - TUITIONFEE_IN
```

**Step 6:** Create a new variable called `TOTAL_IN` which contains the total
amount of money made from tuition per year.

```{r}
#| label: add-total-tuition

colleges_clean6 <- colleges_clean |> 
    select(TOTAL_IN = UGDS x TUITIONFEE_IN)
```

<!-- Swap roles -- Developer becomes Coder, Coder becomes Developer! -->

**Step 7:** Remove every row with missing data.

```{r}
#| label: drop-na

colleges_clean7 <- colleges_clean |> 
  drop.na()
```

At this point we've made seven different datasets, but notice each of these steps
started with `colleges_clean <- colleges_clean |>`. That is pretty redundant! 
Instead, we could have performed all these tasks as one long "pipeline."

**Step 8:** Combine your (fixed) code chunks into a **single** code chunk that
carries out all of the steps necessary to clean the data. 

*Think about coding efficiency -- you __should not__ have multiple calls to the __same__ function!*

```{r}
#| label: all-steps-one-pipeline
# Code combining ALL of your previous steps into ONE pipeline

colleges_clean <- colleges |>
  
```

# Part Two: Identify the Mystery College

Wow! Your best friend Margaret has been accepted to her top choice college!
Unfortunately, Margaret is a very mysterious person, and she won't tell you
directly which college this is. You'll have to use her clues to figure out which
school is her dream school.

**Clues:**

<!-- Swap roles -- Developer becomes Coder, Coder becomes Developer! -->

1.  This college is located in Region 4.

2.  This college's admission rate is below the median rate for the region.

3.  This college **does** charge the same for in- and out-of-state tuition.

4.  The average SAT score of this college is an even number.

<!-- Swap roles -- Developer becomes Coder, Coder becomes Developer! -->

5.  This college is in the state where gooey butter cake, the Chiefs, and the
Anheuser-Busch Clydesdales are a thing.

6.  More than 1,000 undergraduates attend this college.

7.  The total amount of money made from tuition per year for this college is 
greater than or equal to $50,000,000.

8.  Of the three options remaining at this step, Margaret will attend the
college that has the lowest admission rate.

```{r}
#| label: identify-college

```

Submit the college Margaret will attend to the Canvas Quiz.
