---
title: "Untitled"
---


## Random Babies Simulation

Perhaps you have seen the [Random Babies applet](https://www.rossmanchance.com/applets/2021/randombabies/RandomBabies.html)? Suppose one night at a hospital four mothers each give birth to a baby. The hospital is not very organized and looses track of which mother gave birth to each baby, and essentially returns the babies to the mothers at random. Here, we are interested in the number of babies that are correctly returned to their respective mothers. 

**1. Simulate the distribution of the number of babies that are correctly returned. Use 10000 simulations.**

::: callout-tip
Write a function to accomplish one simulation, then use `map_df` to save results
out.
:::

```{r}
randomBabies <- function(nBabies){
  # fill out the function body
}

# map_df code

```

**2. Create a barplot showing the simulated distribution (proportion).**

::: callout-tip
Useful code: 
```{r}
geom_bar(mapping = aes(y = after_stat(count) / sum(after_stat(count))
                       )
         )
```
:::

```{r}
# code for 2
```
